Linear response theory for core set identification

### 2012-05-24 writedown/linear.response/langevin.tex
improve

The core set plays a very important role in the MSM theory. One way to
identify core set is to consider the ``stability'' of the invariant
measure $f_0$ under some mapping $\mt_\alpha$ of time step $\alpha$:
\begin{align}
  C_\alpha = \big\{
  \vect x \,\vert\, \mt_\alpha f_0(\vect x) > f_0(\vect x)
  \big\}
\end{align}
Generally, it is impossible and not necessary to consider the core set
in the $3N$ dimensional configuration space. For example, a
macromolecule solving in water, the DOFs of water is not of interest
at all. Therefore, we consider some phase space compressing function
$\vect y = \vect y(\vect x)$, with $\vect y \in \Omega \subset \mathbb
R^M,\ M\ll 3N$. We call $\Omega$ the observation space.  The projected
invariant measure on the observation space is therefore
\begin{align}
  \mathcal P f_0 (\vect y)
  =
  \int
  \dd \vect x\,
  f_0(\vect x) \,
  \delta (\vect y - \vect y(\vect x)) 
\end{align}
The core set on the observation space is
\begin{align}
  \widetilde
  C_\alpha = \big\{
  \vect y \,\vert\,
  \mathcal P\mt_\alpha f_0(\vect y) > \mathcal P f_0(\vect y)
  \big\}
\end{align}
% with $\widetilde \mt_\alpha$ defined by
% $\widetilde\mt_\alpha = \mathcal P\circ\mt_\alpha$, namely
% \begin{align}
%   \widetilde\mt_\alpha\, p_0(\vect y)
%   =
%   \mathcal P\circ\mt_\alpha f_0 (\vect y)
%   =
%   \int
%   \dd \vect x\,
%   [\mt_\alpha f_0](\vect x) \,
%   \delta (\vect y - \vect y(\vect x)) 
% \end{align}
For convenience, we define $\Delta \mt_\alpha =
\mt_\alpha - \mathcal I$, where $\mathcal I$ is the
identity. If $[\mathcal P\circ\Delta \mt_\alpha f_0](\vect y) > 0$, then
$\vect y$ is in the core set $\widetilde C_\alpha $.
\begin{align}\nonumber
  [\mathcal P\circ\Delta \mt_\alpha f_0](\vect y)
  &=
  \int
  \dd \vect x\,
  \Delta \mt_\alpha f_0(\vect x) \,
  \delta (\vect y - \vect y(\vect x))  \\\nonumber
  &=
  \lim_{\sigma \rightarrow 0}
  \int
  \dd \vect x\,
  \Delta \mt_\alpha f_0(\vect x) \,
  K_{\sigma} (\vect y - \vect y(\vect x))  \\\nonumber
  &=
  \lim_{\sigma \rightarrow 0}
  \int
  \dd \vect x\,
  \Delta \mt_\alpha f_0(\vect x) \,
  K_{\sigma, \vect y} (\vect x)  \\\nonumber
  &=
  \lim_{\sigma \rightarrow 0}
  \int
  \dd \vect x\,
  f_0(\vect x) \,
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
  &=
  \lim_{\sigma \rightarrow 0}\,
  \langle
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
  \rangle_0
\end{align}
where $[\Delta \mt_\alpha]^\dagger$ is the adjoint operator of $\Delta
\mt_\alpha$, which is actually $\Delta \mt_{-\alpha}$. The delta
function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
which is infinitely smooth, compact supported of size $\sigma$,
bounded and $\int
K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
can consider the expectation value of the stability of the mollifier
$K_{\sigma, \vect y}$ instead of the stability of the invariant measure.
This provides the convenience of using the linear response theory showed
in the previous section. By Eqn.~\eqref{eqn:lr},
\begin{align}\nonumber
  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y)
  &=
  \lim_{\sigma \rightarrow 0}\,
  \langle
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
  \rangle_t \\
  &=
  \lim_{\sigma \rightarrow 0}\,
  \Big\{
  \langle
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
  \rangle_0  -
  \beta
  \int_0^t\dd s\,
  F_e(t-s)
  \langle
  j(0)\,
  \mathbb E_{\vect x}\{
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(s)\,
  \}
  \rangle_0
  \Big\}
\end{align}
Therefore, the core set at time $t$ is defined by
\begin{align}
  \widetilde C_\alpha^t =
  \Big\{
  \vect y
  \,\Big\vert\,
  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y) > 0
  \Big\}
\end{align}

>>>>

The core set plays a very important role in the MSM theory. One way to
identify core set is to consider the ``stability'' of the invariant
measure $f_0$ under a propagator $\mt_\alpha$
of time step $\alpha$, which is associated with a
stochasitc process denoted by $\vect Y_\alpha$:
  C_\alpha = \big\{
  \vect x \,\vert\, \mt_\alpha f_0(\vect x) > f_0(\vect x)
  \big\}
Generally, it is impossible and not necessary to consider the core set
in the $3N$ dimensional configuration space. For example, a
macromolecule solving in water, the DOFs of water is not of interest
at all. Therefore, we consider some phase space compressing function
$\vect y = \vect y(\vect x)$, with $\vect y \in \Omega \subset \mathbb
R^M,\ M\ll 3N$. We call $\Omega$ the observation space.  The projected
invariant measure on the observation space is therefore
  \mathcal P f_0 (\vect y)
  \int
  \dd \vect x\,
  f_0(\vect x) \,
  \delta (\vect y - \vect y(\vect x)) 
The core set on the observation space is
  \widetilde
  C_\alpha = \big\{
  \vect y \,\vert\,
  \mathcal P\mt_\alpha f_0(\vect y) > \mathcal P f_0(\vect y)
  \big\}
% with $\widetilde \mt_\alpha$ defined by
% $\widetilde\mt_\alpha = \mathcal P\circ\mt_\alpha$, namely
% \begin{align}
%   \widetilde\mt_\alpha\, p_0(\vect y)
%   =
%   \mathcal P\circ\mt_\alpha f_0 (\vect y)
%   =
%   \int
%   \dd \vect x\,
%   [\mt_\alpha f_0](\vect x) \,
%   \delta (\vect y - \vect y(\vect x)) 
% \end{align}
For convenience, we define $\Delta \mt_\alpha =
\mt_\alpha - \mathcal I$, where $\mathcal I$ is the
identity. If $[\mathcal P\circ\Delta \mt_\alpha f_0](\vect y) > 0$, then
$\vect y$ is in the core set $\widetilde C_\alpha $.
  [\mathcal P\circ\Delta \mt_\alpha f_0](\vect y)
  \int
  \dd \vect x\,
  \Delta \mt_\alpha f_0(\vect x) \,
  \delta (\vect y - \vect y(\vect x))  \\\nonumber
  \lim_{\sigma \rightarrow 0}
  \int
  \dd \vect x\,
  \Delta \mt_\alpha f_0(\vect x) \,
  K_{\sigma} (\vect y - \vect y(\vect x))  \\\nonumber
  \lim_{\sigma \rightarrow 0}
  \int
  \dd \vect x\,
  \Delta \mt_\alpha f_0(\vect x) \,
  K_{\sigma, \vect y} (\vect x)  \\\nonumber
  \lim_{\sigma \rightarrow 0}
  \int
  \dd \vect x\,
  f_0(\vect x) \,
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
  \lim_{\sigma \rightarrow 0}\,
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
where $[\Delta \mt_\alpha]^\dagger$ is the adjoint operator of $\Delta
\mt_\alpha$. The delta
function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
which is infinitely smooth, compact supported of size $\sigma$,
bounded and $\int
K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
can consider the expectation value of the stability of the mollifier
$K_{\sigma, \vect y}$ instead of the stability of the invariant measure.
This provides the convenience of using the linear response theory showed
in the previous section. By Eqn.~\eqref{eqn:lr},
  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y)
  \lim_{\sigma \rightarrow 0}\,
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
  \rangle_t \\\nonumber
  \lim_{\sigma \rightarrow 0}\,
  \Big\{
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
  \rangle_0  -
  \mathbb E_{\vect x}\{
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(\vect X_s)\,
  \}
  \Big\} \\
  \lim_{\sigma \rightarrow 0}\,
  \Big\{
  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
  \rangle_0  -
  \Big\langle
  \mathbb E_{\vect x}
  \big[
  \mathbb E_{\vect X_s}
  (K_{\sigma, \vect y}(\vect Y_\alpha))
  K_{\sigma, \vect y}(\vect X_s)
  \big]
  \Big\rangle_0
  \Big\}
with initial condistions:
  \vect X_0 = \vect x; \quad \vect Y_0 = \vect X_s.
Therefore, the core set at time $t$ is defined by
  \widetilde C_\alpha^t =
  \Big\{
  \vect y
  \,\Big\vert\,
  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y) > 0
  \Big\}
